\bf Visual diagnostics for constrained optimisation with application to guided tours

### 2020-11-15 paper/paper.Rmd
Editing second section

The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general. Examples of those include the existance of multiple maxima (local and global), the tradeof between computational burden and proximity to the maxima,  dealing with noisy objective functions, that might be non smooth and non diferentiable (REF). Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.

@@

Optimization problems are ubiquitous in many areas of study. While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem. These numerical methods follow interative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accurary is achieved. Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the tradeoff between desirable accuracy and computational burden, as well as dealing with noisy objective functions. In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.

In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form. That is, our problem consists of finding the mininum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.  In this scenario, gradient descent methods have taken the stage for many decades now. In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF). The step size is the key in the sucess of the optimization problem as inadequate step 
sizes might lead to a failure in convergence. In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured. In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is stablished $S_k$ then the objective is to find 
the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line sear). Then move on to the next pont following $x_{k+1} = x_k + \nu_kS_k$
\todo[inline]{lconnection with derivative free methods}
 **From Fletcher's book next paragraph --> clean up** Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method]. This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist [@fletcher2013practical]. A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$. Similarly, for  minimisation we would require a decrease in function value at each step. This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.

>>>>

The optimization challenges encounter in the projection pursuit guided tour problem are common to those of optimization in general. Examples of those include the existance of multiple maxima (local and global), the tradeoff between computational burden and proximity to the maxima,  dealing with noisy objective functions that might be non smooth and non diferentiable (REF). Those  are not unique to this context and therefore the visualization tools and optimization methods presented in this paper can be easily apply to any other optimization problems.

@@

Optimization problems are ubiquitous in many areas of study. While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem. These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved. Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions. In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.

In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form. That is, our problem consists of finding the mininum of a function $f \in L^p$ in the unconstrained $\mathbb{R}$ space.  In this scenario, gradient descent methods have taken the stage for many decades now. In gradient descent methods the goal is to find the mininum of an objective function, and the way to do so, is by taken steps  proportional to the negative gradient of the function at the evaluation point, that is $x_{k+1} = x_k -\alpha_kS_k$ where $\alpha_k$ is the step size or the learning rate (REF). The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence. In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured. In line search methods users are required to provide an initial estimate $x_{1}$ and then for each iteration a search direction is established $S_k$ then the objective is to find the value of $\alpha_k \in \mathbb{R}$ that minimizes $f(x_k + \alpha_kS_k)$ with respect to $\alpha_k$ (this step is called the line sear). Then move on to the next pont following $x_{k+1} = x_k + \nu_kS_k$
\todo[inline]{lconnection with derivative free methods}

 **From Fletcher's book next paragraph --> clean up** Theoretical results have demonstrated the global convergence of the algorithm when the step-size $\alpha_k$ is chosen by minimising  $f(x^{(j)} + \alpha* d^{(j)})$ [@curry1944method]. This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist [@fletcher2013practical]. A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: $f^{(j+1)}> f^{(j)}$. Similarly, for  minimisation we would require a decrease in function value at each step. This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.

### 2020-11-15 paper/paper.tex
Editing second section

Optimization problems are ubiquitous in many areas of study. While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem. These numerical methods follow interative approaches that aim at finding the optimum by progressively improving the current solution until a desirable accurary is achieved. Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the tradeoff between desirable accuracy and computational burden, as well as dealing with noisy objective functions. In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.

In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form. That is, our problem consists of finding the mininum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space. In this scenario, gradient descent methods have taken the stage for many decades now. In gradient descent methods the goal is to find the mininum, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF). The step size is the key in the sucess of the optimization problem as inadequate step
sizes might lead to a failure in convergence. In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured. In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is stablished \(S_k\) then the objective is to find
the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line sear). Then move on to the next pont following \(x_{k+1} = x_k + \nu_kS_k\)
\todo[inline]{lconnection with derivative free methods}
\textbf{From Fletcher's book next paragraph --\textgreater{} clean up} Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}. This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist \citep{fletcher2013practical}. A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\). Similarly, for minimisation we would require a decrease in function value at each step. This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.

>>>>

Optimization problems are ubiquitous in many areas of study. While in some cases, analytical solutions can be found, the majority of problems rely on numerical methods to find the optimal solution to a problem. These numerical methods follow interactive approaches that aim at finding the optimum by progressively improving the current solution until a desirable accuracy is achieved. Although that principle seems uncomplicated, optimization methods need to deal with many challenges such as the existance of multiple maxima (local and global), the trade off between desirable accuracy and computational burden, as well as dealing with noisy objective functions. In addition, optimization results might depend on the algorithm starting values affecting the consistency of results.

In this paper we focus on unconstrained optimization as defined in the introduction section and assume that it is not possible to find an analytical solution in closed form. That is, our problem consists of finding the mininum of a function \(f \in L^p\) in the unconstrained \(\mathbb{R}\) space. In this scenario, gradient descent methods have taken the stage for many decades now. In gradient descent methods the goal is to find the mininum of an objective function, and the way to do so, is by taken steps proportional to the negative gradient of the function at the evaluation point, that is \(x_{k+1} = x_k -\alpha_kS_k\) where \(\alpha_k\) is the step size or the learning rate (REF). The step size is the key in the success of the optimization problem as inadequate step sizes might lead to a failure in convergence. In order to solve this problem line minimization or line search methods were proposed (REF) where the step size is chosen in such a way that convergence is ensured. In line search methods users are required to provide an initial estimate \(x_{1}\) and then for each iteration a search direction is established \(S_k\) then the objective is to find the value of \(\alpha_k \in \mathbb{R}\) that minimizes \(f(x_k + \alpha_kS_k)\) with respect to \(\alpha_k\) (this step is called the line sear). Then move on to the next pont following \(x_{k+1} = x_k + \nu_kS_k\)
\todo[inline]{lconnection with derivative free methods}

\textbf{From Fletcher's book next paragraph --\textgreater{} clean up} Theoretical results have demonstrated the global convergence of the algorithm when the step-size \(\alpha_k\) is chosen by minimising \(f(x^{(j)} + \alpha* d^{(j)})\) \citep{curry1944method}. This is however rarely implemented in practice, either because of the high computational demand or because the solution does not exist \citep{fletcher2013practical}. A modified approach, in the case of maximisation, is to impose a mandatory increase in the objective function for each iteration: \(f^{(j+1)}> f^{(j)}\). Similarly, for minimisation we would require a decrease in function value at each step. This approach turns out to be efficient in practical problems, but global convergence cannot be guaranteed in this case.
